\sout{A smart street lighting system based on \rev{an} IPv6-enabled wireless sensor network} \rev{A routing protocol for an urban IoT based on a smart streetlighting wireless sensor network}

### 2014-12-17 2014_ELSEVIER_CEE_SI/paper_cee_special_issue.tex
alterações sobre protocolo IP

Recently, the 6LoWPAN protocol has been standardized by IETF as an adaptation layer to enable IEEE 802.15.4-based low power and lossy networks (LLNs), such as wireless sensor networks, to reuse \rev{the} existing IPv6 protocol for \rev{the} networking layer \citep{RFC4944}. The IP protocol is an open, mature and field proven technology, with free and well documented standards (RFCs), and thus it is very reasonable to reuse it for wireless sensor networking instead of creating new proprietary protocols. However, 6LoWPAN protocol does not define the routing protocol, which shall be defined apart. Thus, IETF has standardized RPL as the routing protocol for IP-based LLNs \citep{RFC6550}.

>>>>

Recently, the 6LoWPAN protocol has been standardized by the IETF \rev{(Internet Engineering Task Force)} as an adaptation layer to enable IEEE 802.15.4-based low power and lossy networks (LLNs), such as wireless sensor networks, to reuse \rev{the} existing IPv6 protocol for \rev{the} networking layer \citep{RFC4944}. \rev{Also, the IETF has standardized RPL as the routing protocol for IP-based LLNs \citep{RFC6550}}. \rev{According to the literature, the main reasons to use IPv6 protocol for the LLNs include the connectivity through the internet to enable the IoT, the 128-bit addressing space that enables uniquely addressable devices (``things''), the network layered architecture and the interoperability \citep{IPv6_LLN_2010, IPv6_LLN_2011, Smart_City_IOT_2014b}.}

\sout{The IP protocol is an open, mature and field proven technology, with free and well documented standards know as RFCs \rev{(Request for Comments)}, and thus it is very reasonable to reuse it for wireless sensor networking instead of creating new proprietary protocols.} \sout{However, 6LoWPAN protocol does not define the routing protocol, which shall be defined apart. Thus, IETF has standardized RPL as the routing protocol for IP-based LLNs}.
